Tree height, microhabitat, and hydraulic traits shape drought responses in a temperate broadleaf forest

### 2019-10-06 manuscript/climate_sensitivity_draft.tex
Tried updating symbols

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Tree size and hydraulic traits shape drought responses in a temperate broadleaf forest},
            pdfauthor={Ian McGregor, Ryan Helcoski, Norbert Kunert, Alan Tepley, Valentine Herrmann, Joseph Zailaa, Atticus Stovall, Neil Pederson, Lawren Sack, Krista Anderson-Teixeira},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

@@

  \title{Tree size and hydraulic traits shape drought responses in a temperate
broadleaf forest}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Ian McGregor, Ryan Helcoski, Norbert Kunert, Alan Tepley, Valentine
Herrmann, Joseph Zailaa, Atticus Stovall, Neil Pederson, Lawren Sack,
Krista Anderson-Teixeira}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

@@

\usepackage{float}
\usepackage{booktabs}
\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{caption}
\captionsetup[table]{font=small}
\captionsetup[figure]{font=small}
\captionsetup[table]{labelformat=empty}
\usepackage{dcolumn}

@@

\subsubsection{Summary}\label{summary}

Predicting forest responses to drought is an increasingly critical task
under climate change effects. Part of the problem is due to the lack of
studies analyzing the confluence of leaf hydraulic traits with
biophysical parameters. In this study, we analyze the interaction
between these two trait groups using forest census data from a 25.6-ha
ForestGEO plot in Virginia (USA). Drought periods were defined by both
Palmer Drought Severity Indices (PDSI) and their identification from
tree-ring records for 12 species representing 97\% of woody
productivity. Each drought scenario (1964-66, 1977, 1999), along with
the overall trend, was then tested against leaf hyraulic trait
measurements and microhabitat biophysical data. Individual-level growth
responses to the three individual droughts were stronger among taller
trees in dominant canopy positions, those in wetter microsites, and for
more drought-sensitive species as assessed by leaf traits (turgor loss
at less negative leaf water potential, greater shrinkage with leaf
dehydration), with substantial variation in the best predictor variables
across given droughts. We conclude that when droughts occur, large
dominant trees, drought-sensitive species, and individuals in wetter
microhabitats are likely to be most strongly affected.

@@

\subsubsection{Introduction}\label{introduction}

Understanding how and why trees respond to drought is critical to
predicting forest drought responses and climate change feedbacks. (1
paragraph on this- I have lots of content on this and will add later.)

{[}Understanding forest responses to drought requires increased
functional understanding of the factors that confer individual-level
vulnerability or resistance.{]} Forests are diverse in terms of tree
sizes and functional traits, and it is known that trees varying in size
and functional traits respond differently to drought (e.g.,
\citep{bennett_larger_2015}; REFS). Therefore, in order to understand
whole-forest response to drought, we need a functional understanding of
how responses vary by tree size, microhabitat, and species. There are 3
fundamental questions that must be addressed:

\emph{First, what drives the observed tendency for large trees to suffer
more during drought?}\\
\citet{bennett_larger_2015} showed that in forests globally, large trees
suffer greater growth reductions during drought. However, this analysis
quantified tree size based on DBH, which has no direct mechanistic
meaning. This study proposed two major mechanisms (besides insects) for
the observed greater drougth growth reductions of large trees: (1)
inherently greater biophysical challenge of being tall; (2) greater
exposure of the crowns of large trees. Counteracting these effects, (3)
the larger root systems of larger trees may confer an advantage in terms
of allowing greater access to water, but it appears that this effect is
usually insufficient to offset the costs of height and/or crown
exposure.

Canopy trees have lower drought resistance because they are exposed to
higher solar radiation, greater wind speeds, and lower humidity.
Alternatively, the generally supressed status of subcanopy trees may be
insufficient to override the benefits of their buffered environment
during drought.

\emph{Second, how do species' traits - alone and in interaction with
tree size - influence drought response?} Analyzing drought responses on
the species level does not fully explain mechanisms and is not feasible
in diverse forests. The solution is a trait-based approach. Leaf
hydraulic traits hold more promise than more
commonly/traditionally-measured traits such as wood density and specific
leaf area (SLA)
(\href{https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/1365-2435.13229}{Medeiros
et al. 2019}).

Commonly measured traits including wood density (REFS), leaf mass per
area \citep{abrams_adaptations_1990} \citep{guerfel_impacts_2009}, and
ring porosity (Elliot et al. 2015, Friedrichs et al. 2009) have been
linked to drought responses and likely correlated with drought
resistance in this forest. However, we hypothesize that leaf hydraulic
traits such as leaf area shrinkage upon dessication (PLA) and turgor
loss point (TLP), which are emerging as potentially more informative
traits but whose effect on drought resistance has never been tested
(CONFIRM), will prove better predictors.

Moreover, there may be an interaction between traits and tree size. It
is possible that the pattern observed by \citet{bennett_larger_2015}
could be caused by smaller trees being more drought resistant.
Alternatively, larger trees may have more drought-resistant traits as
adpatations to greater biophysical challenges.

\emph{Third, are responses similar or variable across individual drought
years?} Droughts are rarely explicitly defined in ecological studies
\citep{slette_how_2019}, yet no two droughts are the same.

We need to understand the the factors confirming drought vulnerability
or resistance not only for extreme droughts with dramatic impacts on
tree growth and mortality, which tend to dominate the literature
\citep{bennett_larger_2015} (REFS), but also for more modest but
frequent droughts--e.g., those with historical return intervals on the
order of a decade.

Here, we combine tree-ring records covering three droughts (1964-66,
1977, 1999), species functional and hydraulic trait measurments, and
forest census data from a 25.6-ha ForestGEO plot in Virginia (USA) to
test a series of hypotheses and associated specific predictions (Table
1) designed to yield functional understanding of how tree size,
microenrivonment, and species' traits collectively shape drought
responses. First, we focus on the role of tree size and its interaction
with microenvironement. We confirm that, consistent with most forests
globally \citep{bennett_larger_2015}, larger-diameter trees have lower
drought resistance in this forest, which is in an ecoregion represented
by only one study in Bennett et al (2015) (H1.0). We then test
hypotheses designed to disentangle the relative importance of tree
height (H1.1), crown exposure (H1.2), and root water access, which
should be greater for larger trees in dry but not perpetually wet
microsites (H1.3). Second, we focus on the role of species' functional
and hydraulic traits and their interaction with tree height. We
hypothesize that drought resistance will follow predicted and observed
patterns in relation to wood density, specific leaf area, and ring
porosity (P2.1a-c), but that leaf hydraulic traits such as leaf area
shrinkage upon dehydration and turgor loss point will prove better
predictors (P2.1d-e). We then test whether these traits correlate with
tree height (P2.2), potentially driving the observed tendency for taller
trees to suffer more during drought (P2.3). Finally, we focused on
variability among droughts, asking how community resistance varied
across droughts (H3.1) and whether the factors confirming vulnerability
or resistance varied across droughts (H3.2).

\emph{Let's put Table 1 here. We need to give the hypotheses/
predictions here. It seems counterintuitive to put the results right up
front like that, but it actually tends to be a higher-impact style of
writing, and recommended
\href{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005619}{here}.}

\textbf{Table 1. Summary of hypotheses, predictions and results}
\includegraphics[width=5.20833in]{tables_figures/Table1.png}

We count predictions as fully supported (or rejected) when the direction
of response matches (or contradicts) the prediction. H2.3-H3.2 are based
on if models containing species traits in H2.1 had dAIC\textgreater{}=2
relative to the appropriate null model or to any alternative
multivariate model within 2 dAIC. Parentheses indicate that predictions
were partially supported (or rejected). In other words, parentheses
indicate when the direction of response matched (or contradicted) the
prediction in some but not all models with dAIC\textless{}2 relative to
the appropriate null model. With categorical variables such as crown
position, a ``(yes)'' notation describes when the trend matched the
prediction, but it wasn't significant A ``(yes/no) or (''no/yes``)''
indicates tendencies in opposite directions in univariate tests versus
the best full models, respectively. ``(yes)

\subsubsection{Materials and Methods}\label{materials-and-methods}

\emph{Study site} Research was conducted at the 25.6 ha ForestGEO
(Global Earth Observatory) study plot at the Smithsonian Conservation
Biology Institute (SCBI) in Virginia, USA (38°53'36.6``N, 78°
08'43.4''W) \citep{andersonteixeira_ctfs-forestgeo:_2015}. SCBI is
located in the central Appalachian Mountains at the northern edge of
Shenandoah National Park. Elevations range from 273-338m above sea level
\citep{gonzalezakre_patterns_2016} with a topographic relief of 65m
\citep{bourg_initial_2013}. Dominant species include \emph{Liriodendron
tulipifera}, oaks (\emph{Quercus} spp.), and hickories (\emph{Carya}
spp.).

\emph{Data collection and preparation} The SCBI ForestGEO plot was
censused in 2008, 2013, and 2018 following standard ForestGEO protocols,
whereby all free-standing woody stems \textgreater{}= 1cm diameter at
breast height (DBH) were mapped, tagged, measured at DBH, and identified
to species \citep{condit_tropical_1998}. From this census data, we used
measurements of DBH from 2008 to calculate historical DBH (described
below), along with tree location in the plot to determine the
topographic wetness index. Furthermore, we used data for all stems
\textgreater{}= 10cm to analyze functional trait composition relative to
tree height. \emph{I added the last sentence-- not sure if you've
described that below yet. I don't think you need to note here which data
we use, so long as that's described in the portions where you describe
the analysis.}

We analyzed tree-ring data from the twelve species contributing most to
woody aboveground net primary productivity (ANPP), which together
comprised 97\% of study plot ANPP between 2008 and 2013
\citep{helcoski_growing_2019}. Cores were obtained in 2010-2011 or
2016-2017 from a breast height of 1.3m using a 5mm increment borer. In
2010-2011, cores were collected from randomly selected live trees of
species with at least 30 individuals of DBH \textgreater{}= 10cm
\citep{bourg_initial_2013}. In 2016-2017, cores were collected from all
trees found dead in the annual mortality census
\citep{gonzalezakre_patterns_2016}. Cores were sanded, measured, and
cross-dated using standard procedures, as detailed in
\citep{helcoski_growing_2019}. The resulting chronologies have been
published in association with \citet{helcoski_growing_2019}: (ITRDB;
GitHub/Zenodo). \emph{is this a database? Yes. Ryan submitted the data
but I don't think its posted yet. We should also cite GitHub/Zenodo
here. I'll come back to that. }

Height measurements (n=\# trees) were taken by several researchers
between 2012 to 2019, and are archived in a public
\href{https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_dimensions/tree_heights}{GitHub
repository}. Measurement methods included manual
\citep[NEON]{stovall_assessing_2018}, digital rangefinders
\citep{andersonteixeira_size-related_2015}, and automatic LiDAR
\citep{stovall_terrestrial_2018}. Rangefinders either used the tangent
method (Impulse 200LR, TruPulse 360R) or the sine method (Nikon
ForestryPro) for calculating heights. The associated errors for using
either method were acknowledged \citep{larjavaara_measuring_2013}.
Species-specific height allometries were developed (Table S\# -
\textbf{ADD THIS TABLE TO SI}). For species that didn't have enough
height measurements, heights were calculated from equations derived from
all species in the study.

For each tree, we combined tree-ring records and allometric equations of
height and bark thickness to retroactively calculate DBH and estimate
height for the years 1950-2009. Prior DBH was estimated using the
following equation, using 2008 as the earliest year for having reliable
DBH measurements:

\[ diamYEAR = dbh2008 - 2*(bark.depth2008) - 2*\Sigma(ring.widthYEAR:ring.width2008) + 2*(bark.depthYEAR) \]

Here, \emph{ring.width} was measured from cores. Bark thickness was
estimated from species-specific allometries based on the bark thickness
data of \citep{andersonteixeira_size-related_2015}. Specifically, we
used linear regression equations on log-transformed data to relate bark
thickness to DBH (Table S\#- \textbf{create table to give these
equations in SI}) and then used these to estimate bark thickness based
on DBH.

Crown positions were recorded in the field during the growing season of
2018 following the crown position protocol from
\citep{jennings_assessing_1999}, whereby positions were ranked as
dominant, codominant, intermediate, or suppressed. As there was no way
to retroactively estimate crown position, we assumed that 2018 crown
position was reflective of each tree's position over the past 60 years.
While some trees undoubtedly changed position, an analysis of crown
position relative to height (Fig. XX) and height change since
\emph{1959} indicated that change was likely slow. {[}\textbf{work on
this-- provide details?}{]}

Topographic wetness index (TWI) was calculated using the
\citep{R-dynatopmodel} package in R.

\begin{figure}[htbp]
\centering
\includegraphics[width=5.20833in]{tables_figures/Figure3.png}
\caption{Location of cored trees}
\end{figure}

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Tree height, microhabitat, and hydraulic traits shape drought responses in a temperate broadleaf forest},
            pdfauthor={Ian McGregor, Ryan Helcoski, Norbert Kunert, Alan Tepley, Erika Gonzalez-Akre, Valentine Herrmann, Joseph Zailaa, Atticus Stovall, Norman Bourg?, William McShea?, Neil Pederson, Lawren Sack, Kristina Anderson-Teixeira},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

@@

  \title{Tree height, microhabitat, and hydraulic traits shape drought responses
in a temperate broadleaf forest}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Ian McGregor, Ryan Helcoski, Norbert Kunert, Alan Tepley, Erika
Gonzalez-Akre, Valentine Herrmann, Joseph Zailaa, Atticus Stovall,
Norman Bourg?, William McShea?, Neil Pederson, Lawren Sack, Kristina
Anderson-Teixeira}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

@@

\usepackage{float}
\usepackage{booktabs}
\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{caption}
\captionsetup[table]{font=small}
\captionsetup[figure]{font=small}
\captionsetup[table]{labelformat=empty}
\captionsetup[figure]{labelformat=empty}
\usepackage{dcolumn}

@@

\raggedright

\hypertarget{title-page}{%
\subsubsection{Title page}\label{title-page}}

\begin{longtable}[]{@{}llll@{}}
\toprule
\begin{minipage}[b]{0.35\columnwidth}\raggedright
text\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright
word count\strut
\end{minipage} & \begin{minipage}[b]{0.22\columnwidth}\raggedright
other\strut
\end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\raggedright
n\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Total word count (excluding summary, references and legends)\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
currently \textasciitilde{}6000 (strict limit 6,500)\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
No.~of figures\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
2 (both colour)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Summary\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
\#\# (limit 200)\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
No.~of Tables\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
5\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Introduction\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
currently \textasciitilde{}1536\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
No of Supporting Information files\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
\#\#\#.\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Materials and Methods\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
currently \textasciitilde{}1442\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Results\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
currently \textasciitilde{}1437\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Discussion\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
currently \textasciitilde{}1881 (limit 30\% of total (not strict), or
1950 if manuscript reaches word limit)\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Acknowledgements\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
\textasciitilde{}75 unwritten\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\newpage

\hypertarget{summary-needs-work}{%
\subsubsection{\texorpdfstring{Summary (\textbf{needs
work})}{Summary (needs work)}}\label{summary-needs-work}}

\begin{itemize}
\item
  As the climate changes, driving increased drought in many forested
  regions around the world, mechanistic understanding of factors
  conferring drought vulnerability and resistance in trees is
  increasingly important. Yet it remains unclear how tree size and
  species' traits interactively shape tree growth responses across
  droughts.
\item
  In this study, we analyze tree-ring records for 12 species
  representing 97\% of woody productivity in the 25.6-ha ForestGEO plot
  in Virginia (USA) to determine how tree size, microhabitat, and
  species' traits interactively shape drought responses across the three
  strongest droughts over the 60 year period from 1950 and 2009.
\item
  Individual-level growth responses to the three individual droughts
  were stronger in three cases: taller trees in dominant canopy
  positions, trees in wetter microsites, and more drought-sensitive
  species as assessed by leaf traits (turgor loss at less negative leaf
  water potential, greater shrinkage with leaf dehydration). However,
  there was substantial variation in the best predictor variables across
  given droughts.
\item
  We conclude that when droughts occur, large dominant trees,
  drought-sensitive species, and individuals in wetter microhabitats
  tend to be most strongly affected.
\end{itemize}

@@

\emph{Key words}: canopy position; drought; Forest Global Earth
Observatory (ForestGEO); hydraulic traits; temperate broadleaf deciduous
forest; tree growth; tree height; tree-ring {[}5-8{]} \emph{Five to
eight key words (in alphabetical order) . Words that are in the title
can, and should, be among these. Very short phrases and scientific names
with their common equivalents (e.g.~Nicotiana tabacum (tobacco)) are
acceptable.}

\newpage

\hypertarget{introduction}{%
\subsubsection{Introduction}\label{introduction}}

Forests globally play a critical role in climate regulation
(\href{https://science.sciencemag.org/content/320/5882/1444}{Bonan
2008}), yet there remains enormous uncertainty as to how the terrestrial
carbon (C) sink, which is dominated by forests, will respond to climate
change
(\href{https://journals.ametsoc.org/doi/full/10.1175/JCLI3800.1}{Friedlingstein
et al.~2006}). An important aspect of this uncertainty lies in responses
to drought (\textbf{REF}). In many forested regions around the world,
the risk of severe drought is increasing
(\href{https://www.nature.com/articles/nclimate2067}{Trenbert et
al.~2014}), even in conjunction with increasing precipitation
(\href{https://www.cambridge.org/core/books/climate-change-2014-impacts-adaptation-and-vulnerability-part-b-regional-aspects/036A899BD52861D61B0D519C5F2B9334}{IPCC
2014}). Global change-type drought has been affecting forests worldwide
(\href{https://doi.org/10.1016/j.foreco.2009.09.001}{Allen et
al.~2015}), and it is expected that future climate change-driven
droughts will severely impact forests around the world
(\href{https://doi.org/10.1016/j.foreco.2009.09.001}{Allen et al.~2015};
\textbf{REFS}). Larger trees tend to suffer more (e.g.,
\citep{bennett_larger_2015};
\href{https://doi.org/10.1038/s41467-019-12380-6}{Stovall et al.~2019}),
resulting in disproportionate impacts on forest C storage
(\href{https://nph.onlinelibrary.wiley.com/doi/10.1111/nph.14633}{Meakem
et al.~2018}). As a result, forest drought responses stand to strongly
impact forest feedbacks to climate change (\textbf{REFS}), yet accurate
characterization of drought responses remains a modeling challenge
(\textbf{REFS})-- in part because some of the mechanisms underlying
drought responses remain unclear. Understanding forest responses to
drought requires increased functional understanding of how tree size,
microhabitat, and species' traits jointly confer individual-level
vulnerability or resistance, and the extent to which their influence is
consistent across droughts.

\textbf{One fundamental question regarding forest responses to drought
is what drives the observed tendency for large trees to suffer more
during drought.} \citet{bennett_larger_2015} showed that in forests
globally, large trees suffer greater growth reductions during drought,
and numerous subsequent studies have reinforced this finding (Stovall et
al.~2019, Hacket-Pain et al.~2016 (DOI 10.1007/s10342-016-0982-7)
\textbf{REFS}). However, this analysis quantified tree size based on
DBH, which has no direct mechanistic meaning. This study proposed two
major mechanisms--besides the tendency for bark beetles to
preferentially attack larger trees (\emph{Pfeifer, E. M., Hicke, J. A.
\& Meddens, A. J. H. Observations and modeling of aboveground tree
carbon stocks and fluxes following a bark beetle outbreak in the western
United States. Glob. Change Biol. 17, 339--350 (2011).})--for the
observed greater drought growth reductions of large trees. First, taller
trees face greater biophysical challenge of lifting water greater
distances against the effects of gravity and friction (* McDowell, N.
G., Bond, B. J., Hill, L., Ryan, M. G. \& Whitehead, D. in Size and age
related changes in tree structure and function (eds Meinzer, F.C. \&
Niinemets, U.) 255--286 (Springer Publishing, 2011); McDowell, N. G. \&
Allen, C. D. Darcy's law predicts widespread forest mortality under
climate warming. Nat. Clim. Change 5, 669--672 (2015); Ryan, M. G.,
Phillips, N. \& Bond, B. J. The hydraulic limitation hypothesis
revisited. Plant Cell Environ. 29, 367--381 (2006).*), and this may
become a greater liability during drought
(\href{https://www.researchgate.net/publication/26322653_Size-dependent_mortality_in_a_Neotropical_savanna_tree_The_role_of_height-related_adjustments_in_hydraulic_architecture_and_carbon_allocation}{Zhang
et al.~2009}). Second, larger trees may have lower drought resistance
because they are more often in the canopy, where they are exposed to
higher solar radiation, greater wind speeds, lower humidity, and lower
\(CO_2\) concentrations (\textbf{REFS-KAT}). Alternatively, the
generally supressed status of subcanopy trees may be insufficient to
override the benefits of their buffered environment during drought.
Potentially counteracting the biophysical challenges faced by large
trees, their larger root systems may confer an advantage in terms of
allowing greater access to water (\textbf{REFS?}); however, it appears
that this effect is usually insufficient to offset the costs of height
and/or crown exposure \citep{bennett_larger_2015}. A final mechanism
that could mediate tree size-related responses to drought is how
hydraulic traits are distributed with respect to size
(\href{https://nph.onlinelibrary.wiley.com/doi/10.1111/nph.14633}{Meakem
et al.~2018}). It is possible that the pattern observed by
\citet{bennett_larger_2015} could be caused if the larger size classes
were dominated by species less adapted to handle drought, be it through
avoidance, resistance, or resiliance. Alternatively, larger size classes
may be dominated by species that are better adapted to inherently
greater biophysical challenges--as is the case in tropical moist forests
of Panama, where larger size classes contain greater proportions of
deciduous species
(\href{https://doi-org.smithsonian.idm.oclc.org/10.2307/3236572}{Condit
et al.~2000};
\href{https://nph.onlinelibrary.wiley.com/doi/10.1111/nph.14633}{Meakem
et al.~2018}). Understanding the mechanisms underlying the tendency for
larger trees to suffer more during drought will require sorting out the
interactive effects of height, canopy position, root water acess, and
species' traits.

\textbf{A second fundamental question regarding forest responses to
drought is how species' traits -- alone and in interaction with tree
size -- influence drought response.} To link drought responss to
fundamental physiological characteristics, and because measuring and
modeling drought responses of every species is infeasible in diverse
forests, it is important to understand how traits shape drought
responses. Commonly measured traits including wood density (\(WD\))
(\textbf{REFS}), leaf mass per area (\(LMA\))
\citep{abrams_adaptations_1990, guerfel_impacts_2009}, and xylem
architecture \citep{kannenberg_linking_2019}(Elliot et al.~2015,
Friedrichs et al.~2009) have been linked to drought responses in
temperate deciduous forests--as well as in other forest biomes
(\textbf{REFS}). However, these traits have less direct linkage to plant
hydraulic function than leaf hydraulic traits such as leaf area
shrinkage upon dessication (\(PLA_{dry}\); \emph{Scoffoni et al.~2013-
DOI: 10.1104/pp.113.221424}) and turgor loss point
(\(\pi_{tlp}\))--i.e., the water potential at which leaf wilting occurs
(\href{https://www.pnas.org/content/113/46/13098.short}{Bartlett et
al.~2016}), which are emerging as traits with potential to explain
greater variation in plant distribution and function than the more
commonly-measured traits such as \(WD\) and \(LMA\)
(\href{https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/1365-2435.13229}{Medeiros
et al.~2019}). There is also evidence that ecosystems with high
hydraulic trait diversity--but not traits like \(LMA\) and \(WD\)-- show
more modest ecosystem flux variation in response to drought
(\href{https://doi.org/10.1038/s41586-018-0539-7}{Anderegg et al.~2018
(\textbf{but review this pub})}), but the ability of hydraulic traits to
tree performance under drought remains untested. (\textbf{But see
D'Orangeville et al.~2018--except that it may not be a well-constructed
test for traits})

\textbf{A final fundamental question regarding forest responses to
drought is whether tree size and species' traits have similar influence
across droughts, or whether drought variability in factors such as
severity, duration, and timing interact with tree size and traits such
that different components of the community respond differently to
different droughts.} No two droughts are the same, and tree growth
responses vary with drought characteristics such as timing and
atmospheric demand
(\href{https://onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/full/10.1111/gcb.14096}{D'Orangeville
et al.~2018}). However, we are not aware of any studies that compare how
tree size and species' traits mediate growth responses across droughts.
(\textbf{BUT ARE WE MISSING SOMETHING??}) While tree-ring studies
provide long-term records of tree responses to multiple droughts (e.g.,
\citep{lloret_components_2011}; D'Orangeville et al.~2018
\textbf{REFS}), these don't test for differential trait effects across
droughts (D'Orangeville et al.~2018) and generally focus on
species-level responses, which preclude consideration of the roles of
tree size and microenvironment. The ecological studies that have shaped
our understanding of the role of tree size and microenvironment in
forest drought responses generally examine only a single drought and
tend to focus disproportionately on extreme droughts with dramatic
impacts (\emph{e.g.},
\href{https://doi.org/10.1016/j.foreco.2009.09.001}{Allen et al.~2015};
\citep{bennett_larger_2015}; Stovall; \textbf{MORE REFS}). Thus, our
knowledge of forest responses to more modest but frequent
droughts--e.g., those with historical return intervals on the order of a
decade--remains more limited. While the tendency for larger trees to
suffer more definitely predominates \citep{bennett_larger_2015}, there
are exceptions (e.g., \textbf{REFS}). There is also evidence that the
degree to which larger trees suffer more increases with the severity of
drought conditions (\citep{bennett_larger_2015}; Stovall et al.~2019).
{[}\emph{Are there any studies showing interactions of drought type with
traits?}{]} Thus, while we expect many of the fundamental mechanisms
shaping drought responses to be universal, we have little undertanding
of how tree size and traits interact with drought characteristics to
result in differential responses across droughts. (\textbf{PELTIER})
